Comparing vector-based and ACT-R memory models using large-scale datasets:\protect\\User-customized hashtag and tag prediction on Twitter and StackOverflow

### 2014-09-23 writeups/dissertation/dissertation.tex
foo

The majority of tag-recommendation systems use some measure of word pair associations to compute the likelihood of a tag, given current context.
It is important that this measure scales well enough to handle the large volume of information present in online social media datasets.
Large social media datasets such as StackOverflow and Twitter contain on the order of millions to hundreds of millions of unique word pair associations between terms and tags.
If one wants to capitalize on all of the information present in the database to build the most accurate measure of relatedness between words and tags, then the association measure used must scale to this large size.

@@

As one of the largest tests, \textcite{Farahat2004} used the PMI index based on a 118GB local database of 10 million pages of the Stanford WebBase Project.
So this implementation did not fit completely in RAM, but a spinning-disk database implementation still scaled well enough for reasonably fast retrieval times for $\mathit{PMI}_{ji}$ values.

@@

Unlike StackOverflow, the entire Twitter dataset is not released to the public.
Due to this and the scale of Twitter, subsets of the Twitter dataset were collected.

\subsubsection{Popular Users Subset}

@@

\subsubsection{Popular Hashtags Subset}

@@

\subsubsection{Randomly Sampled Subset}

Since the entire dataset is available for StackOverflow, randomly-sampled posts were used to test the context models rather than subsetting and creating a popular-tags dataset (analogous to Twitter).
If all Twitter data were available, then the models would have been tested on a randomly-sampled dataset for Twitter as well.
However, since this is not the case for Twitter, the popular-hashtags subset was used instead of a randomly-sampled subset for that site,
while models for StackOverflow were tested on the more challenging dataset where posts were randomly sampled across all posts created on the site.

@@

For each randomly-sampled post for StackOverflow, the prior activation is based on that user's prior tagging history, and not on any global prior of specific tag frequency across users.
This was done so that the prior component used for StackOverflow for both the popular-users and randomly-sampled subsets was the same.

>>>>

The majority of tag-recommendation systems use some measure of word pair associations to compute the likelihood of a tag, given current context.
As a applied concern, it is important that this measure scales well enough to handle the large volume of information present in online social media datasets.
Large social media datasets such as StackOverflow and Twitter contain on the order of millions to hundreds of millions of unique word pair associations between terms and tags.
If one wants to capitalize on all of the information present in the database to build the most accurate measure of relatedness between words and tags, then the association measure used must scale to this large size.

However, scale is also a theoretical concern.
The declarative memory system is tasked to quickly return a single chunk of information that is stored alongside millions to billions of other unique chunks given only the current context and past history of use.
To do this task, it seems plausible that a model of declarative memory must use a representation that produces accurate retrievals even when large amounts of observations are included.
Further, it seems plausible that the time to compute an activation value by a computer is correlated with the time to compute an activation value by the brain.
So any model of declarative memory should also be reasonably computationally efficient even with large co-occurrence datasets.

@@

As one of the largest tests, \textcite{Farahat2004} used the PMI index based on a 118GB local database of 10 million pages of the Stanford WebBase Project.
Even at this large dataset size, the ACT-R declarative memory theory still scaled well enough for reasonably fast retrieval times for $\mathit{PMI}_{ji}$ values.

@@

Unlike StackOverflow, the entire Twitter dataset is not released to the public.
Due to this and the scale of Twitter, only the part of the Twitter dataset relevant to this research was collected.

\subsubsection{Popular Users Dataset}

@@

\subsubsection{Popular Hashtags Dataset}

@@

\subsubsection{Randomly Sampled Dataset}

Since the entire dataset is available for StackOverflow, randomly-sampled posts were used to test the context models rather than creating a popular-tags dataset (analogous to Twitter).
If all Twitter data were available, then the models would have been tested on a randomly-sampled dataset for Twitter as well.
However, since this is not the case for Twitter, the popular-hashtags dataset was used instead of a randomly-sampled dataset for that site,
while models for StackOverflow were tested on the more challenging dataset where posts were randomly sampled across all posts created on the site.

@@

For each randomly-sampled post for StackOverflow, the prior activation is based on that user's prior tagging history, and not on any global prior of specific tag frequency across users.
This was done so that the prior component used for StackOverflow for both the popular-users and randomly-sampled datasets was the same.
